### 2021-02-12 Modules/stabilityEnhancement_freqPairApproach.tex
Sent as draft No.2

\subsection*{Intuition}
Bearing in mind that the system's phase alignment sensitivity resides in \eqref{eq_generalH} via the  term $\exp\Brack{j\dPhi}=\exp\Brack{j\omega\Delta\tau_{pd}}$ and that the round-trip delay ($\tau_{pd}$) cannot be controlled, one may suggest to use lower frequencies. 
Unfortunately, aiming for practical range estimation errors, the transmission of such low frequencies is physically unfeasible. 
Instead, we suggest simultaneous transmission of several frequencies in order to resolve the range error sensitivity. In the following, we suggest a dual frequency (DF) waveform, utilizing two harmonics, $\omega_1$ and $\omega_2$.

\subsection*{Suggested processing scheme}
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.9\linewidth, 
        % grid, 
        tics=10,trim={0 0 0 0}]{./Media/SpatialIIR_APP.png}
            \put (12.5, 64){$\text{FB}_{\vAlphaI{1},\vBetaI{1}}$}
            \put (61, 64){$\text{FB}_{\vAlphaI{2},\vBetaI{2}}$}
            \put (4.5, 59){$z_{1}\rBrace{t}$}
            \put (54, 59){$z_{2}\rBrace{t}$}
            \put (12.9, 41.5){$\text{BPF}_{\omega_{1}}$}
            \put (61.65, 41.5){$\text{BPF}_{\omega_{2}}$}
            \put (6, 51.5){+}
            \put (55, 51.5){+}
            \put (16, 51.5){\footnotesize{$\text{n}_{1}\rBrace{t}$}}
            \put (64.75, 51.5){\footnotesize{$\text{n}_{2}\rBrace{t}$}}
            \put (24.5, 59){\footnotesize{$\Tx_{1}\rBrace{t}$}}
            \put (73.5, 59){\footnotesize{$\Tx_{2}\rBrace{t}$}}
            \put (36.25, 64){\scriptsize{$s_{1}\rBrace{t}$}}
            \put (43, 64){\scriptsize{$s_{2}\rBrace{t}$}}
            \put (18.25, 13){\footnotesize{Harmonic mean}}
            \put (32, 2){$Z_{\text{DF}}$}
            \put (54.75, 24){$\Sigma$}
            \put (21.75, 25){\footnotesize{$\mathcal{F}_{\omega_{1}}$}}
            \put (30.75, 25){\footnotesize{$\mathcal{F}_{\omega_{2}}$}}
        \end{overpic}
    \end{center}
    \caption{Dual-frequency beamformer, consisting of two independent FB blocks and narrowband bandpass filters. The blocks marked by $\mathcal{F}_{\omega_{i}}$ compute the Fourier coefficients in $\omega_{i}$ and their outputs feed the harmonic mean calculator, which generates the DF beamformer's output.}
    \label{fig_app}
\end{figure}
In Fig.~\ref{fig_app}, we demonstrate the use of two independently configured FB instances (see also Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), where each instance is designed to treat a specific frequency band.
A bandpass filter $\text{BPF}_{\omega_{i}},\;i=1,2$ filters a narrowband slice around $\omega_{i}=2\pi f_i$. These filters are used to generate both the transmitted feedback signal ($\text{Tx}_{i}$) and the outputs $z_{i}$. 
The inputs $s_{i}(t) = \exp\Brack{j\omega_{i}t}$  are the two narrowband stimuli signals and $\text{n}_{i}(t)$ describe the additive noise.

Note that in the suggested architecture we do not add array elements, but merely double the beamformer processing effort. 
\par For each FB block, its output is given by
\[
z_{i}(t)=H_{\vBetaI{i},\vAlphaI{i}}\rBrace{\omega_{i}}\exp{\rBrace{j\omega_{i}t}}\ \ {i\in\cBrace{1,2}},
\]
where $\vBetaI{i},\vAlphaI{i}$ are the coefficients of the $i$'th beamformer.
Motivated by the desire to mitigate the $\Delta\phi$ dependency of the system, which appears in the denominator of \eqref{eq_generalH}, we compute the reciprocal of each frequency response, average, and compute the reciprocal again. This leads to the harmonic mean of both beamformers' outputs (see Fig.~\ref{fig_app}), which formally, up to a constant term, takes the form
\begin{equation}
    \label{eqn_HDF_def}
    Z_{\text{DF}} = \abs{H^{-1}_{\vBetaI{1}\vAlphaI{1}}\rBrace{\omega_{1}}+H^{-1}_{\vBetaI{2},\vAlphaI{2}}\rBrace{\omega_{2}}}^{-1}.
\end{equation}
% Note that, as in IIR filters, the system stabilizes after a certain period, therefore the Fourier weights should be calculated after stabilization.
% As a rule of thumb, simulations show that the stabilization period is around $100R/c$. 
\ifdefined\useOmega
Straight forward derivations, Allowing $g$ to be frequency dependent, leads to
\else
For convenience, we use subscripts instead of formal $\omega$ dependency such that $\phiI{i}\triangleq\phi\rBrace{\omega_{i}}, \gI{i}\triangleq{}g\rBrace{\omega_{i}}, \vdI{i}\triangleq\vecnot{d}\rBrace{\omega_{i}}$.
Also, $r_i\triangleq{}\gI{i}/\gIHat{i}$ is denoted to be the gain mismatch at $\omega_i$.

@@

\begin{proof}
See App.~\ref{apdx_thrm_DF}. 
\end{proof}
\par Assuming close frequencies, $Z_{\text{DF}}$ in  \eqref{eqn_H_DF_general} closely resembles the single frequency (SF) beampattern in  \eqref{eqn:GeneralFeedbackTransferFunction}, where the range related phase $\phi$ is now replaced by  $\phiI{1}-\phiI{2}=(\omega_1-\omega_2)\tau_{pd}$. 
Hence, one may significantly mitigate the range mismatch distortion of the beampattern by selecting close stimuli frequencies.
% Also, $\tau_n$ is arbitrary, hence the discussion is not restricted to a specific array geometry.
We also note that throughout the development of \eqref{eqn_H_DF_general}, we did not assume any specific array geometry, hence this result is valid for arbitrary arrays and not just ULA.
\subsection*{Numerical example}
Consider a radio frequency carrier of $10$~GHz and typical range error of $\Delta{}R_{\text{rt}}=10$~m, which is $333\frac{1}{3}\lambda$ (assuming  speed of light, $c=3\cdot 10^{8}$~m/s). The single frequency beampattern distortion, being periodic in $\lambda$, will closely resemble the $0.3\lambda$ error plot presented in Fig.~\ref{fig_rangError}. Assume that we aim to achieve a maximal phase error of $\Delta \phi=0.01\pi$. Hence, when using DF architecture, the dictated frequency separation must satisfy
\begin{equation}\label{eq_DF_phErr}
\abs{(\omega_1-\omega_2)\frac{\Delta{}R_{\text{rt}}}{c}}<0.01\pi.
\end{equation}
or equivalently, for maximal range error of $10$~m, a frequency separation of
\[
\abs{f_1-f_2}<0.005 c/\Delta{}R_{\text{rt}}=150 \,\text{kHz}
\]
is required.

\subsection*{Dual frequency simulation}

@@

    \resizebox{.89\linewidth}{!}{

>>>>

\subsection{Intuition}
Bearing in mind that the system's phase alignment sensitivity resides in \eqref{eq_generalH} via the  term $\exp\Brack{j\dPhi}=\exp\Brack{j\omega\Delta\tau_{pd}}$ and that the round-trip delay ($\tau_{pd}$) cannot be controlled, one may suggest to use lower frequencies. 
Unfortunately, aiming for practical range estimation errors, the transmission of such low frequencies is physically unfeasible.
As the medium for the wave's propagation in the context of localization is not free space, physical considerations are to be taken when choosing the frequency band of use.
Following the thorough spectrum use overview in \cite{barclay2003propagation}, among the notable considerations for choosing frequency in localization related schemes are
\begin{itemize}
    \item \textbf{Propagation efficiency}\\
    Due to the chemical structure of the atmosphere, some frequencies are propagated efficiently while other are absorbed \cite{wen2015foundations} by the atmosphere layers, being unsuitable for localization applications such as RADAR.
    \item \textbf{Antenna length}\\
    As described in \cite{barclay2003propagation}, although low ($<30_{\text{Khz}}$) frequencies propagate efficiently in the earth-atmosphere wave guide \cite{wen2015foundations}, the wavelength ($\sim10_{\text{Km}}$), hence the needed antenna length is too large for any practical use.
    \item \textbf{Signal bandwidth}\\
    As the carrier signal's frequency decreases, so does the available bandwidth due to aliasing.
\end{itemize}
Henceforth, inspired by known trigonometrical identities where mathematical of two harmonics produces new frequencies which are the sum (or subtraction) of the original signals, we suggest simultaneous transmission of several frequencies instead of the unfeasible option of transmitting low frequencies. 
In the following we suggest a dual frequency (DF) waveform, utilizing two harmonics, $\omega_1$ and $\omega_2$, and show that range error sensitivity may be resolved.

\subsection{Suggested processing scheme}
\begin{figure}[t!]
    \begin{center}
        \begin{overpic}[width=0.65\linewidth, 
        % grid, 
        tics=10,trim={0 0 0 0}]{./Media/SpatialIIR_APP.png}
            \put (12.5, 64){$\text{FB}_{\vAlphaI{1},\vBetaI{1}}$}
            \put (61, 64){$\text{FB}_{\vAlphaI{2},\vBetaI{2}}$}
            \put (4.5, 59){$z_{1}\rBrace{t}$}
            \put (54, 59){$z_{2}\rBrace{t}$}
            \put (12.9, 41.5){$\text{BPF}_{\omega_{1}}$}
            \put (61.65, 41.5){$\text{BPF}_{\omega_{2}}$}
            \put (6, 51.5){+}
            \put (55, 51.5){+}
            \put (16, 51.5){\footnotesize{$\text{n}_{1}\rBrace{t}$}}
            \put (64.75, 51.5){\footnotesize{$\text{n}_{2}\rBrace{t}$}}
            \put (24.5, 59){\footnotesize{$\Tx{}_{1}\rBrace{t}$}}
            \put (73.5, 59){\footnotesize{$\Tx{}_{2}\rBrace{t}$}}
            \put (36.25, 64){\scriptsize{$s_{1}\rBrace{t}$}}
            \put (43, 64){\scriptsize{$s_{2}\rBrace{t}$}}
            \put (18.75, 13){\footnotesize{Harmonic mean}}
            \put (32, 2){$Z_{\text{DF}}$}
            \put (54.75, 24){$\Sigma$}
            \put (22, 25){\footnotesize{$\mathcal{F}_{\omega_{1}}$}}
            \put (30.75, 25){\footnotesize{$\mathcal{F}_{\omega_{2}}$}}
        \end{overpic}
    \end{center}
    \caption{Dual-frequency beamformer, consisting of two independent FB blocks and narrowband bandpass filters. The blocks marked by $\mathcal{F}_{\omega_{i}}$ compute the Fourier coefficients in $\omega_{i}$ and their outputs feed the harmonic mean calculator, which generates the DF beamformer's output.}
    \label{fig_app}
\end{figure}
In Fig.~\ref{fig_app}, we demonstrate the use of two independently configured FB instances (see also Fig.~\ref{fig:Proposed_spatialIIR_ARCH}), where each instance is designed to treat a specific frequency band.
A bandpass filter $\text{BPF}_{\omega_{i}},\;i=1,2$ filters a narrowband slice around $\omega_{i}=2\pi f_i$. These filters are used to generate both the transmitted feedback signal ($\text{Tx}_{i}$) and the outputs $z_{i}$. 
The inputs $s_{i}(t) = \exp\Brack{j\omega_{i}t}$  are the two narrowband stimuli signals and $\text{n}_{i}(t)$ represents the additive noise.

Note that in the suggested architecture we do not add array elements, but merely double the beamformer processing effort. 
\par For each FB block, its output is given by
\[
z_{i}(t)=H_{\vBetaI{i},\vAlphaI{i}}\rBrace{\omega_{i}}\exp{\rBrace{j\omega_{i}t}}\ \ {i\in\cBrace{1,2}},
\]
where $\vBetaI{i},\vAlphaI{i}$ are the coefficients of the $i$'th beamformer.
Motivated by the desire to mitigate the $\Delta\phi$ dependency of the system, which appears in the denominator of \eqref{eq_generalH}'s RHS, we compute the reciprocal of each frequency response, average, and compute the reciprocal again. This leads to the harmonic mean of both beamformers' outputs (see Fig.~\ref{fig_app}), which formally, up to a constant term, takes the form
\begin{equation}
    \label{eqn_HDF_def}
    Z_{\text{DF}} = \abs{H^{-1}_{\vBetaI{1}\vAlphaI{1}}\rBrace{\omega_{1}}+H^{-1}_{\vBetaI{2},\vAlphaI{2}}\rBrace{\omega_{2}}}^{-1}.
\end{equation}
% Note that, as in IIR filters, the system stabilizes after a certain period, therefore the Fourier weights should be calculated after stabilization.
% As a rule of thumb, simulations show that the stabilization period is around $100R/c$. 
For convenience, we use subscripts instead of formal $\omega$ dependency such that $\phiI{i}\triangleq\phi\rBrace{\omega_{i}}, \gI{i}\triangleq{}g\rBrace{\omega_{i}}, \vdI{i}\triangleq\vecnot{d}\rBrace{\omega_{i}}$.
Also, $r_i\triangleq{}\gI{i}/\gIHat{i}$ is denoted to be the gain mismatch at $\omega_i$.

@@

\begin{proof}
See App.~\ref{apdx_thrm_DF}. 
\end{proof}
\par Assuming close frequencies, $Z_{\text{DF}}$ in  \eqref{eqn_H_DF_general} closely resembles the single frequency (SF) beampattern in  \eqref{eqn:GeneralFeedbackTransferFunction}, where the range related phase $\phi$ is now replaced by  $\phiI{1}-\phiI{2}=(\omega_1-\omega_2)\tau_{pd}$. 
Hence, one may significantly mitigate the range mismatch distortion of the beampattern by selecting close stimuli frequencies.
% Also, $\tau_n$ is arbitrary, hence the discussion is not restricted to a specific array geometry.
We also note that throughout the development of \eqref{eqn_H_DF_general}, we did not assume any specific array geometry, hence this result is valid for arbitrary arrays and not just ULA.
\subsection{Numerical example}
Consider a radio frequency carrier of $10$~GHz and typical range error of $\Delta{}R_{\text{rt}}=10$~m, which is $333\frac{1}{3}\lambda$ (assuming  speed of light, $c=3\cdot 10^{8}$~m/s). The single frequency beampattern distortion, being periodic in $\lambda$, will closely resemble the $0.3\lambda$ error plot presented in Fig.~\ref{fig_rangError}. Assume that we aim to achieve a maximal phase error of $\Delta \phi=0.01\pi$. Hence, when using the DF architecture, the dictated frequency separation must satisfy
\begin{equation}\label{eq_DF_phErr}
\abs{(\omega_1-\omega_2)\frac{\Delta{}R_{\text{rt}}}{c}}<0.01\pi.
\end{equation}
or equivalently, for a maximal range error of $10$~m, a frequency separation of
\[
\abs{f_1-f_2}<0.005 c/\Delta{}R_{\text{rt}}=150 \,\text{kHz}
\]
is required.

\subsection{Dual frequency simulation}

@@

    %\resizebox{.89\linewidth}{!}
    {

### 2021-02-12 front/abbrevs.tex
Sent as draft No.2

\abbreviationsAndNotation{
\begin{tabular}{p{2cm}@{:\quad}l}
QED & quod erat demonstratum \\
$c$ & the speed of light \\
$a\pm b$ & the closed interval $\left[a-b,a+b\right]$ \\
\end{tabular}
}

>>>>

\abbreviationsAndNotation{
\begin{longtable}{p{2cm}@{:\quad}l}
DOA & Direction of arrival \\
RADAR & RAdio Detection And Ranging \\
FIR & Finite impulse response \\
IIR & Infinite impulse response \\
ULA & Uniform linear array \\
FIM & Fisher information matrix \\
SNR & Signal to noise ratio \\
MUSIC & Multiple signal classification (algorithm)\\
RHS & Right hand side \\
FB & Feedback beamformer \\
SF & Single frequency \\
DF & Dual frequency \\
HPBW & Half power beamwidth \\
PDF & Probability density function \\
CRLB & Cram\'er Rao lower bound \\
CRB & Cram\'er Rao bound \\
PCA & Principal component analysis \\
ICA & Independent components analysis \\
MVDR & Minimum Variance Distortionless Response \\
UWB & Ultra wide band \\
HOS & High order statistics \\
$\vecnot{v}$ & A vector \\
$v_{i}$ & The $i$-th element of vector $\vecnot{v}$ \\
$\vecnot{v}^{T}$ & Transpose of vector $\vecnot{v}$ \\
$\vecnot{v}^{*}$ & Conjugate of vector $\vecnot{v}$ \\
$\vecnot{v}^{H}$ & Hermitian (conjugated) transpose of vector $\vecnot{v}$ \\
$\vecnot{A}$ & A Matrix \\
$A_{i,j}$ & The $i,j$-th element of matrix $\vecnot{A}$ \\
$\vecnot{A}^{T}$ & Transpose of matrix $\vecnot{A}$ \\
$\vecnot{A}^{*}$ & Conjugate of matrix $\vecnot{A}$ \\
$\vecnot{A}^{H}$ & Hermitian (conjugated) transpose of matrix $\vecnot{A}$ \\
$N$ & Number of elements in array \\
$d$ & Array's inter-element spacing \\
$\omega$ & Radial frequency \\
$c$ & Wave's propagation velocity \\
$\delta{}x$ & Small diff in $x$ \\
$\lambda$ & Wavelength \\
$t$ & Time \\
$\theta_{d}$ & Geometrical angle \\
$\psi$, $\theta$ & Electrical phase \\
\vecnot{d} & Steering vector \\
$B$ & beampattern \\
$R$ & Distance to target \\
$z\rBrace{t}$ & Array's temporal output \\
$Z\rBrace{\omega}$ & Array's frequency-domain output \\
$\mathcal{D}$ & Directivity \\
$\mathcal{H}$ & Normalized beampattern \\
\end{longtable}
}

### 2021-02-12 main/intro.tex
Sent as draft No.2

The array size and the number of its elements ($N$) has significant influence on the obtained array performance, such as SNR improvement, spatial separation capabilities and its spatial response's degrees of freedom (DOF). For example, the influential MUSIC algorithm \cite{schmidt1986multiple} enables the localization of signals arriving from up to $N-1$ distinctive directions of arrival (DOA), by projecting the array manifold onto the noise subspace.

@@

Using a similar approach, the $2q$-MUSIC algorithm \cite{chevalier2006high}, enables the use of $N^{2q}$ ``virtual elements'', by calculating the $q$'th order statistics.

@@

The main advantage of the first (i.e. virtual array) over the latter (i.e. minimum redundancy) is the obtainable analytical expression for the improvement in spatial selectivity when using virtual arrays.
In the minimum redundancy approach no such expression is available.
On the other side, the computation effort is considerably higher when the 2-nd order statistics are computed.

@@

Adaptive processing schemes \cite{frost1972algorithm,manolakis2000statistical}, being a wide and active research area, were also suggested trying to adaptively estimate and suppress the noise component in impinging signals by minimization of the receiver's output energy with some constraints.

@@

\par Pursuing other approaches to improve the array's spatial performance, ULA spatial array processing analogy to finite impulse response (FIR) \cite{van1988beamforming} and the infinite impulse response (IIR) superior performance (e.g. narrower transition regions and higher sidelobes attenuation) gave rise to the question ``what are the equivalent spatial domain processing methods which will be analogous to temporal IIR filtering?''

@@

The first one was to estimate the time of arrival (TOA) difference between two consecutive sensors and to synthetically generate the recursive part of the IIR filter, entirely in the time-domain.
The second approach suggested to consider overlapping subsets of one large ULA as finite approximation to an infinite array. 
The former approach heavily relies on the accuracy of the delay estimation which are amplified due to the fact that both the desired signal and the errors are continuously fed back into the filter.
Close inspection of the latter approach reveals that even though the sub-arrays are each weighted separately, each sensor's output is eventually linearly weighted when taking into account all related sub-arrays.

@@

Also, ultra-wideband (UWB) filters, which sample spatial snapshots of the signal and recursively process it in temporal domain were designed in \cite{Hum2009BeamformingFilters}, using the $2D$ spatio-temporal plane wave representation as a straight line angled according to the DOA.
\par
TODO: Elaborate on the 2/3D planar from (by order)
\begin{itemize}
    \item Highly Selective Three-Dimensional Recursive Beam Filters Using Intersecting Resonant Planes 
    \item Three-Dimensional Cone Filter Banks
    \item A Speed-Optimized Systolic Array ProcessorArchitecture for Spatio-Temporal 2-D IIRBroadband Beam Filters
    \item UWB Beamforming Using 2-D Beam Digital Filters
\end{itemize}

@@

% \par The outline of this paper is as follows. We first formulate the classic spatial beamforming setup in Sec.~\ref{sec:setup}. Then, in Sec.~\ref{sec_introduceFeedback}, we propose our novel feedback-based architecture, and formulate its spatial response.
% Searching for localization performance maximization, Sec.~ \ref{sec_FIM} discusses the information-theory related considerations for the array configuration, utilizing the Fisher Information Matrix (FIM) in the context of the target's range and DOA estimation.
% In Sec.~\ref{sec_Performance}, we evaluate some key features of the proposed beamforming with feedback. Specifically, we compute the array beamwidth, its peak-to-sidelobe ratio and the array directivity, showing significant improvement compared to traditional beamforming without spatial feedback. 
% In Sec.~\ref{sec_app}, we simulate the proposed processing scheme, and emphasize its sensitivity to range errors. We then suggest a strategy which mitigates this sensitivity. Finally, concluding remarks are given in Sec.~\ref{sec_conclusions}.

>>>>

The array size and the number of its elements ($N$) has significant influence on the obtained array performance, such as SNR improvement, spatial separation capabilities and its spatial response's degrees of freedom (DOF).
\par Early algorithms for DOA estimation were based on the beamforming concept \cite{krim1996two}, where the array's reception beam is steered (mechanically or electronically) to multiple directions in search of received energy peaks which are interpreted as valid emitters.
One notable improvement to the convectional beamformer \cite{van2004optimum} was Capon's beamformer  (aka MVDR in acoustics related literature) which attempts to minimize the power contributed by noise and any signals coming from other directions than the DOA of interest, while maintaining a fixed gain
in the "look direction".
\par The most practically used approach in DOA estimators, referred as the sub-space based estimation, involves invariant sub-spaces analysis of observed covariance matrices.
Early research include PCA and errors-in-variables time series analysis \cite{krim1996two}.   
However, the tremendous interest in the subspace approach is mainly due to the introduction of the
MUSIC (Multiple SIgnal Classification) algorithm \cite{schmidt1986multiple} which decomposes the estimated covariance matrix to its signal and noise eigenvectors, using orthogonal projection assuming that the noise space is orthogonal to the array manifold.
It is interesting to note that while earlier works were mostly derived in the context of time series analysis and later applied to the sensor array problem, MUSIC was indeed originally presented as a DOA estimator.
\par A Less practical, but optimal in the sense of RMSE \cite{krim1996two}, is the statistical based ML approach which tries to find the most probable DOA assuming a known PDF.
Though this approach served as a basis for extensive research, leading to more advanced concepts such as the ICA \cite{hyvarinen1999survey}, in the context of practical DOA estimation it's computational effort is high.
Adding the fact that sub-optimal methods such as MUSIC were proven achieve the CRB under some practical assumptions (e.g. non-coherent signals) \cite{stoica1989music}, such methods (i.e. the sub optimal estimators) are substantially more common in practice. 

@@

\par Using a similar approach, the $2q$-MUSIC algorithm \cite{chevalier2006high}, enables the use of $N^{2q}$ ``virtual elements'', by calculating the $q$'th order statistics.

@@

% The main advantage of the first (i.e. virtual array) over the latter (i.e. minimum redundancy) is the obtainable analytical expression for the improvement in spatial selectivity when using virtual arrays.
% In the minimum redundancy approach no such expression is available.
% On the other side, the computation effort is considerably higher when the 2-nd order statistics are computed.

@@

\par Adaptive processing schemes \cite{frost1972algorithm,manolakis2000statistical}, being a wide and active research area, were also suggested trying to adaptively estimate and suppress the noise component in impinging signals by minimization of the receiver's output energy with some constraints.

@@

\par In \cite{van1988beamforming}, ULA based beamforming and temporal FIR filtering are shown to be mathematically analogous, where the DOA acts as the spatial version of temporal frequency.
Inspired by this analogy, which is thoroughly discussed in Sec.~\ref{sec:prlm_FIR_IIR}, we were motivated to find the spatial counterpart of the temporal IIR filter, both from academic curiousness and the known advantages of the IIR filter's frequency (DOA by analogy) response, hoping to gain spatial performance by finding the right architecture.
Therefore we formulated a question, ``what are the equivalent spatial domain processing methods which will be analogous to temporal IIR filtering?'' which served as a guide throughout our research.

@@

The first one (see Fig.~\ref{fig_intro_wen1}) was to estimate the time of arrival (TOA) difference between two consecutive sensors and to synthetically generate the recursive part of the IIR filter, entirely in the time-domain.
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/WEN_method1.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{Synthetic delay emulated for the recursive part of the spatial IIR filter.}
    \label{fig_intro_wen1}
\end{figure}
The second approach suggested to consider overlapping subsets of one large ULA as finite approximation to an infinite array (see Fig.~\ref{fig_intro_wen2}).
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/WEN_method2.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{Referring to consecutive sub arrays' output as sequential output of the same ULA.}
    \label{fig_intro_wen2}
\end{figure}
Obviously, the former approach is very sensitive to the synthetic delay accuracy and involves temporal domain processing which we try to avoid in this work to achieve purely spatial processing.  
Also, close inspection of the latter approach reveals that even though the sub-arrays are each weighted separately, each sensor's output is eventually linearly weighted when taking into account all related sub-arrays.

@@

\par Other interesting works \cite{Hum2009BeamformingFilters,madanayake2008speed} in the field of ultra-wideband (UWB) treat the spatial and temporal diversity, which reside in multiple consecutive snapshots of the array output, as if the signal is sampled in two independent temporal frequencies.
Inspired by \cite{bruton1985three}, the authors show that impinging plane waves are represented as straight lines tilted according to their DOA and suggest methods to design line filters in the spatio-temporal plane.
In \cite{Hum2009BeamformingFilters}, the authors also address the fact that ideal straight line filter may not be achieved due to the fact that the filter is periodic in both frequencies, causing the 2D frequency response the ``bend`` close to the sampling frequencies (see Fig.~\ref{fig_intro_UWB_bend}).
\begin{figure}[ht!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/UWB_bend.png}
            % \put (50, 62.5) {\footnotesize{$r=0$}}
        \end{overpic}
    \end{center}
     \caption{2D spatio-temporal frequency response (dB) of practical UWB filter. The black line is angled according to the desired DOA.}
    \label{fig_intro_UWB_bend}
\end{figure}

@@

\par In this work, we use the following conventions; vectors and matrices are denoted by lower and upper case bold italic letters respectively. 
$v_{i}$ and $A_{ij}$ are the $i$-th element of the vector $\vecnot{v}$ and the $i,j$-th element of the matrix $\vecnot{A}$.
Also, $\vecnot{A}^{T}$, $\vecnot{A}^{*}$ and $\vecnot{A}^{H}$ are the transpose, conjugate and Hermitian transpose of the matrix $\vecnot{A}$ respectively.
\par In the followings, we elaborate in Chapter~\ref{chap:prelims} on some relevant theoretic concepts which serve as the foundation of this presentation.
Then our novel architecture is presented and analysed in Chapter.~\ref{chap:firstchap} followed by a thorough concluding discussion on the findings and future research in Chapter.~\ref{chap:futureResearch}.

### 2021-02-12 main/prelims.tex
1. Intro exapnsion - WIP 2. Move appdxs to end of thesis - DONE
Before implementing Tsvi's comments
Sent as draft No.2

\chapter{Preliminaries}
\label{chap:prelims}

As a preliminary discussion to the presentation of out contribution, we overview some the key theoretical basis of the thesis.
In Sec.~\ref{sec:prlm_propWaveField}, the wave propagation, being the most basic concept for array processing, is discussed.
It is then followed by a short overview of array-processing in Sec.~\ref{sec:prlm_sensorArrays}.
To accommodate the discussion for spatial filtering, we also mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}.
As an additional theoretical concept, arriving from the information theory, which will be used in the main chapter (Ch.~\ref{chap:firstchap}), is the Fisher Information Matrix, covered in Sec.~\ref{sec:prlm_FIM}.
\par 
For an easier reading treat each section as independent, where notations and symbols are each defined in the specific section and are not mutual, nor used in the main chapter.

% \begin{definition}
% The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
% \end{definition}

\section{Propagating Wave Fields}
\label{sec:prlm_propWaveField}
\input{Modules/sec_prlm_propWaveField}
% \section{Digital Filter Design - Finite/Infinite Response Filters}
% \label{sec:prlm_FILTERS}
% \input{Modules/sec_prlm_FILTERS}
\section{Sensor arrays}
\label{sec:prlm_sensorArrays}
\input{Modules/sec_prlm_sensorArrays}
\section{Time-Space Signals}
\label{sec:prlm_timeSpaceSig}
\input{Modules/sec_prlm_timeSpaceSig}
% \section{Sensor array performance measure}
% \label{sec:prlm_sensorArrayPerf}
% \input{Modules/sec_prlm_arrPerf}
\section{Fisher Information Matrix}
\label{sec:prlm_FIM}
\input{Modules/sec_prlm_FIM}

>>>>

\chapter{Preliminaries}
\label{chap:prelims}
As a preliminary discussion to the presentation of our contribution, we overview some the key theoretical basis of the thesis.
First, a short overview of array-processing is presented in Sec.~\ref{sec:prlm_sensorArrays}, accompanied by a complementary wave propagation discussion in Sec.~\ref{sec:prlm_propWaveField}.
To accommodate the discussion for spatial filtering, we also mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}.
In Sec.~\ref{sec:prlm_localization}, we further focus the spatial filtering discussion to localization and overview some of its fundamentals.
Following the FIR/IIR filters related discussion in Chapter.~\ref{chap:intro}, we further elaborate and compare the two architectures in Sec.~\ref{sec:prlm_FIR_IIR}.
To complete the basics overview, we also discuss array performance measuring in Sec.~\ref{sec:prlm_array_perf} following the classical localization related literature.
% \begin{definition}
% The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
% \end{definition}

% \section{Digital Filter Design - Finite/Infinite Response Filters}
% \label{sec:prlm_FILTERS}
% \input{Modules/sec_prlm_FILTERS}
\section{Sensor arrays}
\label{sec:prlm_sensorArrays}
\input{Modules/sec_prlm_sensorArrays}
\section{Propagating Wave Fields}
\label{sec:prlm_propWaveField}
\input{Modules/sec_prlm_propWaveField}
\section{Spatio-temporal processing}
\label{sec:prlm_timeSpaceSig}
\input{Modules/sec_prlm_timeSpaceSig}
\section{Localization}
\label{sec:prlm_localization}
\input{Modules/sec_prlm_localization}
% \section{Sensor array performance measure}
% \label{sec:prlm_sensorArrayPerf}
% \input{Modules/sec_prlm_arrPerf}
\section{The analogy between array processing and temporal digital filtering}
\label{sec:prlm_FIR_IIR}
\input{Modules/sec_prlm_FIR_IIR}
\section{Array spatial performance}
\label{sec:prlm_array_perf}
\input{Modules/sec_prlm_array_perf}
